---
title: "Assignment 5: Data Visualization"
author: "David Amanfu"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on
Data Visualization

## Directions

1.  Change "Student Name" on line 3 (above) with your name.
2.  Work through the steps, **creating code and output** that fulfill
    each instruction.
3.  Be sure to **answer the questions** in this assignment document.
4.  When you have completed the assignment, **Knit** the text and code
    into a single PDF file.
5.  After Knitting, submit the completed exercise (PDF file) to the
    dropbox in Sakai. Add your last name into the file name (e.g.,
    "Fay_A05_DataVisualization.Rmd") prior to submission.

The completed exercise is due on Monday, February 14 at 7:00 pm.

## Set up your session

1.  Set up your session. Verify your working directory and load the
    tidyverse and cowplot packages. Upload the NTL-LTER processed data
    files for nutrients and chemistry/physics for Peter and Paul Lakes
    (use the tidy
    \[`NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv`\]
    version) and the processed data file for the Niwot Ridge litter
    dataset (use the \[`NEON_NIWO_Litter_mass_trap_Processed.csv`\]
    version).

2.  Make sure R is reading dates as date format; if not change the
    format to date.

```{r, setup}
#1a
library(tidyverse)
library(cowplot)
library(ggplot2)
getwd()
knitr::opts_knit$set(root.dir = "~/Desktop/Duke MPP/Environ Data /Environmental_Data_Analytics_2022/")
```

```{r}
#1b
getwd()
NTL_LTER_Lake_Nutrients <- read.csv("Data/Processed/NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv", stringsAsFactors = TRUE)
NTL_LTER_Lake_ChemistryPhysics <- read.csv("Data/Processed/NTL-LTER_Lake_ChemistryPhysics_PeterPaul_Processed.csv", stringsAsFactors = TRUE)
NEON_NIWO_Litter <- read.csv("Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv", stringsAsFactors = TRUE)

#2 
class(NEON_NIWO_Litter$collectDate)
class(NTL_LTER_Lake_ChemistryPhysics$sampledate)
class(NTL_LTER_Lake_Nutrients$sampledate)
NTL_LTER_Lake_ChemistryPhysics$sampledate <- as.Date.factor(NTL_LTER_Lake_ChemistryPhysics$sampledate, tryFormats ="%Y-%m-%d")
NTL_LTER_Lake_Nutrients$sampledate <- as.Date(NTL_LTER_Lake_Nutrients$sampledate, tryFormats ="%Y-%m-%d")
NEON_NIWO_Litter$collectDate <- as.Date(NEON_NIWO_Litter$collectDate, tryFormats ="%Y-%m-%d")
class(NEON_NIWO_Litter$collectDate)
class(NTL_LTER_Lake_ChemistryPhysics$sampledate)
class(NTL_LTER_Lake_Nutrients$sampledate)

```

## Define your theme

3.  Build a theme and set it as your default theme.

```{r, fig.height = 3, fig.width = 4, warning = FALSE}
#3
library(ggthemes)
library(hrbrthemes)
library(extrafont)
#remotes::install_version("Rttf2pt1", version = "1.3.8")
library(extrafontdb)
#font_import()
#fonts()

AmanfuTheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "right", legend.direction = "vertical",) #alternative: legend.position + legend.justification
AmanfuTheme2 <- theme_ipsum()+
  theme(legend.position = "bottom",
        legend.key = element_rect(fill = "white", colour = "black"),legend.direction = "horizontal",
        legend.title = element_text(face = "bold"))
AmanfuTheme3 <- theme_fivethirtyeight()+theme(legend.position = "right", legend.direction = "vertical") 
theme_set(AmanfuTheme2)


```

## Create graphs

For numbers 4-7, create ggplot graphs and adjust aesthetics to follow
best practices for data visualization. Ensure your theme, color
palettes, axes, and additional aesthetics are edited accordingly.

4.  \[NTL-LTER\] Plot total phosphorus (`tp_ug`) by phosphate (`po4`),
    with separate aesthetics for Peter and Paul lakes. Add a line of
    best fit and color it black. Adjust your axes to hide extreme values
    (hint: change the limits using `xlim()` and `ylim()`).

```{r}
#4
tppo4 <-ggplot(NTL_LTER_Lake_Nutrients,aes(x= tp_ug, y=po4)) + 
  geom_point(aes(shape = lakename, color = lakename, fill = lakename)) + 
  ylim(0,50) + 
  labs(title = "Total Phosphorus v Phosphate", x = "Phosphorus", y= "Phosphate") +
  geom_smooth(method = "lm",color = "black")
tppo4
```

5.  \[NTL-LTER\] Make three separate boxplots of (a) temperature, (b)
    TP, and (c) TN, with month as the x axis and lake as a color
    aesthetic. Then, create a cowplot that combines the three graphs.
    Make sure that only one legend is present and that graph axes are
    aligned.

```{r}
#5
tempplot <- ggplot(NTL_LTER_Lake_Nutrients,aes(x=factor(month))) + 
  geom_boxplot(aes(y=temperature_C,color=lakename)) + 
  labs(title = "Temperature (C) by Month", x = "Month", y= "Temperature (C)")
tp_plot <- ggplot(NTL_LTER_Lake_Nutrients,aes(x= factor(month))) + 
  geom_boxplot(aes(y=tp_ug, color = lakename)) +
  #geom_violin(aes(y=tp_ug,color=lakename))+
  labs(title = "Total Phosphorus", x = "Month", y= "Total Phosphorus")
tn_plot <- ggplot(NTL_LTER_Lake_Nutrients,aes(x= factor(month))) +   
  geom_boxplot(aes(y=tn_ug, color = lakename)) +
  labs(title = "Total Nitrates", x = "Month", y= "Total Nitrates")

tempplot
tp_plot
tn_plot

temptptn <- plot_grid(
  tempplot+theme(legend.position = "none",plot.margin=margin(10,30,0,30), plot.title = element_text(size=11)),
  tp_plot+theme(legend.position = "none", plot.margin=margin(10,30,0,30), plot.title = element_text(size=12)),
  tn_plot+theme(legend.position = "none", plot.margin=margin(10,30,10,30), plot.title = element_text(size=12)),
  get_legend(tempplot+theme(legend.direction = "horizontal")),
  align="hv",nrow = 4,rel_heights = c(1.5,1.5,1.5,.25),axis="tb")
temptptn
```

Question: What do you observe about the variables of interest over
seasons and between lakes?

> Answer: So it appears that there's a lot more variation in
> Phosophorous and Nitrates(?) in Peter lake than there is in Paul lake
> over the course of the seasons. It is also notable that phosphorous
> levels seem to follow temperature patterns, while nitrates do not.
> there's nothing particularly interesting or notable about the
> temperature plot; it appears that these lakes are either next to each
> other or at worst in similar climates. We've also got a lot of
> potential outliers in our measurements, which I'm not exactly sure
> what that might suggest -- perhaps too small a sample size? Or perhaps
> too largely bucketed/clustered x values. Perhaps weekly might give us
> smoother data.

6.  \[Niwot Ridge\] Plot a subset of the litter dataset by displaying
    only the "Needles" functional group. Plot the dry mass of needle
    litter by date and separate by NLCD class with a color aesthetic.
    (no need to adjust the name of each land use)

7.  \[Niwot Ridge\] Now, plot the same plot but with NLCD classes
    separated into three facets rather than separated by color.

```{r}
#6
Niwot_needle <- ggplot(subset(NEON_NIWO_Litter, functionalGroup=="Needles"))+
  geom_point(aes(y=dryMass,x=collectDate, color = nlcdClass, shape=nlcdClass))+
  theme(legend.position = "right",legend.direction = "vertical")
Niwot_needle
#7
Niwot_needle_facet <- ggplot(subset(NEON_NIWO_Litter, functionalGroup=="Needles"))+
  geom_point(aes(y=dryMass,x=collectDate,shape=nlcdClass))+
  facet_wrap(vars(nlcdClass),nrow=3)+
  theme(legend.position = "right",legend.direction = "vertical")
Niwot_needle_facet
```

Question: Which of these plots (6 vs. 7) do you think is more effective,
and why?

> Answer: For information that is actually quite clustered, it can
> sometimes help to plot with facets rather than on a single plot. I
> think Plot 7 is more effective in this case because we have better
> clarity on the different classes of needles collected. Plot 6 is super
> crowded and not a lot of information can be gleaned from it.
