---
title: "Assignment 09: Data Scraping"
author: "David Amanfu"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

# Total points:

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on data scraping. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Fay_09_Data_Scraping.Rmd") prior to submission.


## Set up 
1. Set up your session:

* Check your working directory
* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Set your ggplot theme

```{r, initialize}
#1a.
getwd()
knitr::opts_knit$set(root.dir = "~/Desktop/Duke MPP/Environ Data /Environmental_Data_Analytics_2022/")

```


```{r, message = FALSE}
#1b
# install.packages("rvest")
# install.packages("dataRetrieval")
# install.packages("tidycensus")
library(agricolae)
library(corrplot)
library(cowplot)
library(dataRetrieval)
library(extrafont)
library(extrafontdb)
library(ggpubr)
library(ggthemes)
library(hrbrthemes)
library(Kendall)
library(leaflet)
library(lubridate)
library(mapview)
library(rvest)
library(sf)
library(tidycensus)
library(tidyverse)
library(trend)
library(tseries)
library(viridis)
library(zoo)

#Disable on-the-fly projections
sf::sf_use_s2(FALSE)
#Fix Mapview
mapviewOptions(fgb = FALSE)

AmanfuTheme2 <- theme_ipsum()+
  theme(legend.position = "bottom",
        legend.key = element_rect(fill = "white", colour = "black"),legend.direction = "horizontal",
        legend.title = element_text(face = "bold"))
theme_set(AmanfuTheme2)
```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2019 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Change the date from 2020 to 2019 in the upper right corner.
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2020>

Indicate this website as the as the URL to be scraped. (In other words, read the contents into an
`rvest` webpage object.)

```{r set.the.scraping.website}
#2
# https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2020
Durham_LWSP <- read_html('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2017')
Durham_LWSP
```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PSWID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Average Daily Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to three separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third "Municipality", and the last should be a vector of 12 numeric values, with the first value being 36.0100.

```{r scrape.the.data}
#3
water.system.name <- Durham_LWSP %>% html_nodes('table:nth-child(7) tr:nth-child(1) td:nth-child(2)') %>% html_text()
pwsid <- Durham_LWSP %>% html_nodes('tr:nth-child(1) > td:nth-child(5)') %>% html_text()
ownership <- Durham_LWSP %>% html_nodes('table:nth-child(7) tr:nth-child(2) td:nth-child(4)') %>% html_text()

mgd.key <- ':nth-child(32) td:nth-child(9) , :nth-child(32) td:nth-child(6) tr:nth-child(2) :nth-child(9), :nth-child(32) td:nth-child(6), :nth-child(32) td:nth-child(3)'
max.withdrawals.mgd <- Durham_LWSP %>% html_nodes(mgd.key) %>% html_text()

#These failed trials are from using the selector gadget in Safari and Firefox:
# ':nth-child(32) td:nth-child(6) , td:nth-child(9), :nth-child(32) td:nth-child(6), :nth-child(32) td:nth-child(3)'
#  'tr:nth-child(2) td:nth-child(9) , :nth-child(32) td:nth-child(6), :nth-child(32) td:nth-child(3), :nth-child(32) td:nth-child(3)' 
# ':nth-child(31) td:nth-child(9) , tr:nth-child(4) :nth-child(9) tr:nth-child(3) tr:nth-child(2) :nth-child(9), :nth-child(31) td:nth-child(6), :nth-child(31) td:nth-child(3)'

```


4. Convert your scraped data into a dataframe. This dataframe should have a column for each of the 4 variables scraped and a row for the month corresponding to the withdrawal data. Also add a Date column that includes your month and year in data format. (Feel free to add a Year column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data in order. You can overcome this by creating a month column in the same order the data are scraped: Jan, May, Sept, Feb, etc...

5. Plot the max daily withdrawals across the months for 2020

```{r create.a.dataframe.from.scraped.data}
#4

scrapetest <- data.frame("Year"=rep("2017",12),
                         "Month"=c("Jan","May","Sep","Feb","Jun","Oct","Mar","Jul","Nov","Apr","Aug","Dec"),
                         "System"=rep(water.system.name,12),
                         "PWSID"=rep(pwsid,12),
                         "Ownership"=rep(ownership,12),
                         "Withdrawals"=as.double(max.withdrawals.mgd)) %>% mutate("yearchar" =ym(paste0(Year,"-",Month)))
#5
scrapetest
durham2020 <- ggplot(scrapetest,aes(x=yearchar))+geom_point(aes(y=Withdrawals))+labs(title="2020 Max Withdrawals, Durham",x="Date")
durham2020


```

6. Note that the PWSID and the year appear in the web address for the page we scraped. Construct a function using your code above that can scrape data for any PWSID and year for which the NC DEQ has data. **Be sure to modify the code to reflect the year and site scraped**.

```{r construct.a.scraping.function}
#6.

the_facility <- '03-32-010'
the_year <- 2015

scrape.it <- function(the_year, the_facility){
  
  #Retrieve the website contents 
  the_base_url <- 'https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid='
  the_scrape_url <- paste0(the_base_url, the_facility, '&year=', the_year)
  the_website <- read_html(the_scrape_url)

  #Set the element address variables (determined in the previous step)
  water.system.name_node <- 'table:nth-child(7) tr:nth-child(1) td:nth-child(2)'
  pwsid_node <- 'tr:nth-child(1) > td:nth-child(5)'
  ownership_node <- 'table:nth-child(7) tr:nth-child(2) td:nth-child(4)'
  max.withdrawals.mgd_node <- 'th~ td+ td'
    #':nth-child(32) td:nth-child(9) , :nth-child(32) td:nth-child(6) tr:nth-child(2) :nth-child(9), :nth-child(32) td:nth-child(6), :nth-child(32) td:nth-child(3)'
    #':nth-child(31) td:nth-child(9) , tr:nth-child(4) :nth-child(9) tr:nth-child(3) tr:nth-child(2) :nth-child(9), :nth-child(31) td:nth-child(6), :nth-child(31) td:nth-child(3)'
  #Scrape the data items
  water.system.name <- the_website %>% html_nodes(water.system.name_node) %>% html_text()
  pwsid <- the_website %>%   html_nodes(pwsid_node) %>%  html_text()
  ownership <- the_website %>% html_nodes(ownership_node) %>% html_text()
  max.withdrawals.mgd <- the_website %>% html_nodes(max.withdrawals.mgd_node) %>% html_text()
  
  #Construct a dataframe from the scraped data
  df_withdrawals <- data.frame("Year"=rep(the_year,12),
                           "Month"=c("Jan","May","Sep","Feb","Jun","Oct","Mar","Jul","Nov","Apr","Aug","Dec"),
                           "System"=rep(water.system.name,12),
                           "PWSID"=rep(pwsid,12),
                           "Ownership"=rep(ownership,12),
                           "Withdrawals"=as.double(max.withdrawals.mgd)) %>% 
                      mutate("yearchar" =ym(paste0(Year,"-",Month)))
   
  #Pause for a moment - scraping etiquette
  #Sys.sleep(1) #uncomment this if you are doing bulk scraping!
  
  #Return the dataframe
  return(df_withdrawals)
}

```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2015

```{r fetch.and.plot.Durham.2015.data}
#7
scrapetest2015 <-scrape.it(2015,'03-32-010')
scrapetest2015
durham2015 <- ggplot(scrapetest2015,aes(x=yearchar))+geom_point(aes(y=Withdrawals))+labs(title="2015 Max Withdrawals, Durham",x="Date")
durham2015

```

8. Use the function above to extract data for Asheville (PWSID = 01-11-010) in 2015. Combine this data with the Durham data collected above and create a plot that compares the Asheville to Durham's water withdrawals.

```{r fetch.and.plot.Asheville.2015.data}
#8
Asheville2015 <-scrape.it(2015,'01-11-010')
DurhAshe <- union(scrapetest2015,Asheville2015)
DurhAshe
DurhAsheville2015 <- ggplot(DurhAshe,aes(x=yearchar,y=Withdrawals))+geom_point(aes(color=System))+labs(title="2015 Max Withdrawals, Durham & Asheville",x="Date")
DurhAsheville2015

```


9. Use the code & function you created above to plot Asheville's max daily withdrawal by months for the years 2010 thru 2019.Add a smoothed line to the plot.

```{r}
#9
yeargap <- rep(2010:2019)
yeargap
our_facility <- '01-11-010'
the_dfs <- lapply(X = yeargap,
                  FUN = scrape.it,
                  the_facility=our_facility)
Asheville20102019 <- bind_rows(the_dfs)
#Asheville20102019
Asheville1019 <- ggplot(Asheville20102019,aes(x=yearchar,y=Withdrawals))+
                  geom_smooth(method=lm)+geom_point(aes(color=System))+labs(title="2010-2019 Max Withdrawals, Asheville",x="Date")
Asheville1019
```

>Question: Just by looking at the plot (i.e. not running statistics), does Asheville have a trend in water usage over time?
It does! It appears that it is pretty constant from 2010 through 2018, and then we see a marked difference in water usage, given the increase starting in about late 2019.